A Switching Planner for Combined Task and Observation Planning

### 2011-02-05 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/switching.tex
switching is starting to get written again

\subsection{Sequential Planning in POMDPs}

A sequential session uses a classical planner to compute a trace. The
latter encapsulates assumptions about: (1) the true underlying state,
(2) how execution will progress, and (3) the possibility of the agent
eventually holding strong beliefs about the truth values of specific
state propositions. Here we describe the deterministic planning
problem, derived from the DTPDDL model, that admits plans which
correspond to traces.

In a deterministic-action POMDP all the uncertainty in state and
action is expressed in the $(\pp{:init})$ declaration. Our approach
uses the structure of that, as it occurs in the problem description,
to define a set of state-assumptions available to sequential planning.
%%
Writing \#\ if the value of a proposition is unspecified, for
$\pp{DORA}$ we have the following assumptions:

\small
\begin{tabular}{cccc}
\hline
Probability & (is-in DORA)  & (is-in box)  & (is-in cup) \\
\hline
.24 & kitchen & office & office \\
.06 & kitchen & kitchen & office \\
.56 & kitchen & office & kitchen \\
.14 & kitchen & kitchen & kitchen \\
.7 & kitchen & \# &  kitchen\\
.3 & kitchen & \# & office \\
.8 & kitchen & office & \# \\
.2 & kitchen & kitchen & \# \\
1.0 & kitchen & \# & \# \\
\hline
\end{tabular}
\normalsize

\noindent An assumption corresponds to a {\em
relaxed} visitation of the root term of $(\pp{:init})$. In this
relaxed case, a conjunctive term is visited iff its atomic subterms
are visited, and zero or more of its immediate probabilistic subterms
are visited. The starting state, $\state_0$, for sequential planning
is an {\em abstract} state that captures the unique assumption which
has probability $1$. In $\pp{DORA}$, that is:

@@

In order to make assumptions available to the sequential planner, we
add one {\em assumptive} action $\assumptiveS{i}$ to the problem for
each element, $\prob_i (T_i)$, of each probabilistic term from
$(\pp{:init})$. The physics of these actions is as
follows: $\assumptiveS{i}$ can be executed if no $\assumptiveS{j}$,
$j \neq i$, has been executed from the same probabilistic term, and,
either $(\pp{probabilistic}~..\prob_i~(T_i)..)$ is in the root
conjunct, or it occurs in $T_k$ for some executed
$\assumptiveS{k}$. Also, an assumption about a proposition
$\prop$ can only be made once, and must be sequenced before any POMDP
actions that have $\prop$ in their precondition or effect.
%%
In other words, an assumptive action
must be applied to make a truth assignment to \prop, and then that
truth assignment can be modified by the DTPDDL domain operators.
%%
%%
Executing $\assumptiveS{i}$ in $\state$ effects a transition to a
successor state $\state^{T_i}$ with probability $\prob_i$, and
$\state^\bot$ with probability $1 - \prob_i$. Here, $\state^{T_i}$ is
the union of $\state$ with atomic terms from $T_i$ with the
proposition denoting the undefined fact is deleted. For example, if
the action adds $(=(\pp{is-in}~\pp{box})~\pp{office})$, then it must
delete $(=(\pp{is-in}~\pp{box})~\#)$. State $\state^\bot$ is an added
sink, where $\forall\action$ $\reward(\state^\bot, \action) = 0$ and
$\transProb(\state^\bot, \action, \state^\bot) = 1$.
%%
Lastly, we have that if $\action$ has a precondition that is not true
with probability $\theta$, a given threshold, in the underlying belief
$\bstate$, then immediately before it is executed blindly the system
switches to a contingent session (described below).

%% For sequential planning the operators from the POMDP model are
%% available with semantics that accommodate the above abstraction as
%% follows: In the deterministic model a proposition $\prop$ can be
%% thought to have a ternary interpretation at a state \state, as either
%% {\em true}, written $\prop\in\state$, {\em false}, written
%% $\prop\not\in\state$, or {\em unspecified}, with a slight abuse of
%% notation written $\prop\#\in\state$. For example, in $\pp{DORA}$
%% $(=(\pp{is-in}~\pp{box})\pp{office})\#\in\state_0$. For POMDP actions
%% $\action\in\actions$ in the deterministic model, if
%% $\prop \in \poss(\action)$, or if $\prop$ is the subject of a positive
%% or negative effect of $\action$, then $\action$ is not applicable
%% in \state\ if $\prop\#\in\state$.

%% Addressing now the switching semantics of action execution. I
%% %% In detail, we halt the sequential
%% session at \action\ if $\exists
%% \state$ s.t. $\bstate(\state)>0$, 
%% $\prop\in\poss(\action)$ , and $\prop\not\in\state$. According to the
%% semantics of action execution there must be a $\state'$ s.t.
%% $\bstate(\state')>0$ and $\prop\in\state'$. Otherwise, \action\ must
%% not have been scheduled at \bstate.

%% then $\reward(\state,\action)=10$, and
%% otherwise if $\prop\not\in\state'$ then
%% $\reward(\state',\action)=-10$.

We now describe the optimisation criteria given to the classical
planner. Where $\prob_i$ is the probability that the $i^{th}$
sequenced action $\action_i$ does not transition to $\state^\bot$, we
define the value of a trace
$\state_0, \action_0, \state_1, \action_1,.., \state_N$ from the
deterministic model to equal:

@@

\subsection{Decision-Theoretic Planning in Abstractions}

>>>>

\subsection{Sequential Sessions}

We consider only deterministic-action POMDs, where all the uncertainty
in state and action is expressed in the $(\pp{:init})$ declaration. In
sequential sessions, this declaration is used by our approach to
define a starting state, and the set of state-assumptions available to
sequential planning.  Writing \#\ if the value of a proposition is
unspecified, continuing our example from the previous section, we have
the following assumptions:

\small
\begin{tabular}{cccc}
\hline
Probability & (is-in DORA)  & (is-in box)  & (is-in cup) \\
\hline
%% .24 & kitchen & office & office \\
%% .06 & kitchen & kitchen & office \\
%% .56 & kitchen & office & kitchen \\
%% .14 & kitchen & kitchen & kitchen \\
.7 & kitchen & \# &  kitchen\\
.3 & kitchen & \# & office \\
.8 & kitchen & office & \# \\
.2 & kitchen & kitchen & \# \\
1.0 & kitchen & \# & \# \\
\hline
\end{tabular}
\normalsize

\noindent Each assumption corresponds to one distinct {\em
relaxed} visitation of the root term of $(\pp{:init})$. In this
relaxed case, a conjunctive term is visited iff its atomic subterms
are visited, and zero or one of its immediate probabilistic subterms
are visited. For a sequential session, the starting state corresponds
to the unique assumption with probability $1$ -- I.e., the union of
facts from the root conjunctive term. In our example, the starting
state is:

@@

We also annotate DTPDDL model with an assumptive action
$\assumptiveS{i}$ for each element $\prob_i (T_i)$, of each
probabilistic term from $(\pp{:init})$. Here, $\assumptiveS{i}$ can be
executed if no $\assumptiveS{j}$, $j \neq i$, has been executed from
the same probabilistic term, and, either
$(\pp{probabilistic}~..\prob_i~(T_i)..)$ is in the root conjunct, or
it occurs in $T_k$ for some executed $\assumptiveS{k}$. 
%%
We add additional constraints that forbid the scheduling of
assumptions about facts after actions with preconditions, or effects,
that mention those facts. For example, the robot cannot assume it is
plugged into a power source immediately after it unplugs itself.
%%
Executing $\assumptiveS{i}$ in a state $\state$ effects a transition
to a successor state $\state^{T_i}$ with probability $\prob_i$, and
$\state^\bot$ with probability $1 - \prob_i$. Here, $\state^{T_i}$ is
the union of $\state$ with atomic terms from $T_i$. State
$\state^\bot$ is an added sink.

%% %%
%% Lastly, we have that if $\action$ has a precondition that is not true
%% with probability $\theta$, a given threshold, in the underlying belief
%% $\bstate$, then immediately before it is executed blindly the system
%% switches to a contingent session (described below).

We now describe the optimisation criteria used during sequential
sessions. Where $\prob_i$ is the probability that the $i^{th}$
sequenced action $\action_i$ does not transition to $\state^\bot$, we
define the value of a trace
$\state_0, \action_0, \state_1, \action_1,.., \state_N$ from the
deterministic model to equal:

@@

\subsection{DT Sessions}
